Connecting the Popularity Adjusted Block Model to the Generalized Random Dot Product Graph for Clustering and Parameter Estimation

### 2020-10-21 summary.Rmd
ore simulations

Then $P = X Y^\top$ and $Y = X \Pi$ where 
$\Pi = \begin{bmatrix} 
1 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 \\
0 & 0 & 0 & 1 & 0 & 0 & 0 & 0 & 0 \\
0 & 0 & 0 & 0 & 0 & 0 & 1 & 0 & 0 \\
0 & 1 & 0 & 0 & 0 & 0 & 0 & 0 & 0 \\
0 & 0 & 0 & 0 & 1 & 0 & 0 & 0 & 0 \\
0 & 0 & 0 & 0 & 0 & 0 & 0 & 1 & 0 \\
0 & 0 & 1 & 0 & 0 & 0 & 0 & 0 & 0 \\
0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 1
\end{bmatrix}$

@@

@@

```{r clust_err_k, fig.cap = 'IQR of clustering error using Algorithm 2 (red) compared against subspace clustering (blue) for sample sizes from 256 to 4096 and number of clusters from 2 to 4. Simulations were repeated 50 times for each sample size. The $y$-axis is shown on a logarithmic scale.', fig.width = 10}
clustering.df <- readr::read_csv('clustering-k.csv')
# mod.max.df <- readr::read_csv('mod-max-k.csv') %>% 
#   dplyr::group_by(K, n) %>% 
#   dplyr::summarise(med.err.mm = median(error.mm),
#                    first.q.mm = quantile(error.mm, .25),
#                    third.q.mm = quantile(error.mm, .75)) %>% 
#   dplyr::ungroup()
clustering.df %>%
  dplyr::group_by(K, n) %>%
  dplyr::summarise(med.err = median(error),
                   first.q = quantile(error, .25),
                   third.q = quantile(error, .75),
                   med.err.ssc = median(error.ssc),
                   first.q.ssc = quantile(error.ssc, .25),
                   third.q.ssc = quantile(error.ssc, .75)) %>%
  dplyr::ungroup() %>% 
  ggplot() +
  # scale_x_log10() + 
  labs(y = 'community detection error', 
       colour = NULL) +
  geom_line(aes(x = n, y = med.err,
                colour = 'Algorithm 2')) +
  geom_errorbar(aes(x = n, ymin = first.q, ymax = third.q,
                    colour = 'Algorithm 2')) + 
  geom_line(aes(x = n, y = med.err.ssc,
                colour = 'SSC')) + 
  geom_errorbar(aes(x = n, ymin = first.q.ssc, ymax = third.q.ssc,
                    colour = 'SSC')) + 
  # geom_line(data = mod.max.df,
  #           aes(x = n, y = med.err.mm,
  #               colour = 'modularity maximization')) + 
  # geom_errorbar(data = mod.max.df,
  #               aes(x = n, ymin = first.q.mm, ymax = third.q.mm,
  #                   colour = 'modularity maximization')) + 
  scale_colour_brewer(palette = 'Set1') + 
  facet_wrap(~ K, labeller = 'label_both') + 
  scale_x_continuous(breaks = c(256, 512, 1024, 2048, 4096)) +
  scale_y_log10()
```

@@

```{r clust_err_ct_sim, fig.cap = 'IQR of counts of misclustered vertices using Algorithm 2 (red) compared against subspace clustering (blue) for sample sizes from 64 to 2048. Simulations were repeated 100 times for each sample size. The $y$-axis is shown on a logarithmic scale.', cache = FALSE, fig.width = 10}
clustering.df <- readr::read_csv('clustering-k.csv')
clustering.df %>%
  dplyr::group_by(n, K) %>%
  dplyr::summarise(med.err = median(error),
                   first.q = quantile(error, .25),
                   third.q = quantile(error, .75),
                   med.err.ssc = median(error.ssc),
                   first.q.ssc = quantile(error.ssc, .25),
                   third.q.ssc = quantile(error.ssc, .75)) %>%
  dplyr::ungroup() %>% 
  ggplot() +
  # scale_x_log10() + 
  labs(y = 'misclustering count', 
       colour = NULL) +
  geom_line(aes(x = n, y = med.err * n,
                colour = 'Algorithm 2')) +
  geom_errorbar(aes(x = n, ymin = first.q * n, ymax = third.q * n,
                    colour = 'Algorithm 2')) + 
  geom_line(aes(x = n, y = med.err.ssc * n,
                colour = 'SSC')) + 
  geom_errorbar(aes(x = n, ymin = first.q.ssc * n, ymax = third.q.ssc * n,
                    colour = 'SSC')) + 
  scale_colour_brewer(palette = 'Set1') + 
  facet_wrap(~ K, labeller = 'label_both') + 
  scale_y_log10() +
  scale_x_continuous(breaks = c(256, 512, 1024, 2048, 4096))
```

@@

```{r clust_err_k_imba, fig.cap = 'IQR of clustering error using Algorithm 2 (red) compared against subspace clustering (blue) for sample sizes from 256 to 4096 and number of clusters from 2 to 4. Simulations were repeated 50 times for each sample size. The $y$-axis is shown on a logarithmic scale.', fig.width = 10}
clustering.df <- readr::read_csv('clustering-k-imbalanced.csv')
clustering.df %>%
  dplyr::group_by(K, n) %>%
  dplyr::summarise(med.err = median(error),
                   first.q = quantile(error, .25),
                   third.q = quantile(error, .75),
                   med.err.ssc = median(error.ssc),
                   first.q.ssc = quantile(error.ssc, .25),
                   third.q.ssc = quantile(error.ssc, .75)) %>%
  dplyr::ungroup() %>% 
  ggplot() +
  # scale_x_log10() + 
  labs(y = 'community detection error', 
       colour = NULL) +
  geom_line(aes(x = n, y = med.err,
                colour = 'Algorithm 2')) +
  geom_errorbar(aes(x = n, ymin = first.q, ymax = third.q,
                    colour = 'Algorithm 2')) + 
  geom_line(aes(x = n, y = med.err.ssc,
                colour = 'SSC')) + 
  geom_errorbar(aes(x = n, ymin = first.q.ssc, ymax = third.q.ssc,
                    colour = 'SSC')) + 
  scale_colour_brewer(palette = 'Set1') + 
  facet_wrap(~ K, labeller = 'label_both') + 
  scale_x_continuous(breaks = c(256, 512, 1024, 2048, 4096)) +
  scale_y_log10()
```

>>>>

Then $P = X Y^\top$ and $Y = X \Pi$ where 
$\Pi = \begin{bmatrix} 
1 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 \\
0 & 0 & 0 & 1 & 0 & 0 & 0 & 0 & 0 \\
0 & 0 & 0 & 0 & 0 & 0 & 1 & 0 & 0 \\
0 & 1 & 0 & 0 & 0 & 0 & 0 & 0 & 0 \\
0 & 0 & 0 & 0 & 1 & 0 & 0 & 0 & 0 \\
0 & 0 & 0 & 0 & 0 & 0 & 0 & 1 & 0 \\
0 & 0 & 1 & 0 & 0 & 0 & 0 & 0 & 0 \\
0 & 0 & 0 & 0 & 0 & 1 & 0 & 0 & 0 \\
0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 1
\end{bmatrix}$

@@

In these simulations, Algorithm 2 is compared against sparse subspace 
clustering \cite{soltanolkotabi2014} \cite{noroozi2019estimation} and 
modularity maximization \cite{307cbeb9b1be48299388437423d94bf1} \cite{igraph}, 
and Algorithm 4 is compared against an MLE-based estimator 
\cite{noroozi2019estimation} \cite{307cbeb9b1be48299388437423d94bf1}.

@@

```{r clust_err_k, fig.cap = 'IQR of clustering error using Algorithm 2 (red) compared against modularity maximization (blue) and sparse subspace clustering (green) for sample sizes from 256 to 4096 and number of underlying communities from 2 to 4. Simulations were repeated 40 times for each sample size. The $y$-axis is shown on a logarithmic scale.', fig.width = 10}
clustering.df <- readr::read_csv('clustering-k.csv')
clustering.df %>%
  dplyr::group_by(K, n) %>%
  dplyr::summarise(med.err = median(error),
                   first.q = quantile(error, .25),
                   third.q = quantile(error, .75),
                   med.err.ssc = median(error.ssc),
                   first.q.ssc = quantile(error.ssc, .25),
                   third.q.ssc = quantile(error.ssc, .75),
                   med.err.mm = median(error.mm),
                   first.q.mm = quantile(error.mm, .25),
                   third.q.mm = quantile(error.mm, .75)) %>%
  dplyr::ungroup() %>% 
  ggplot() +
  # scale_x_log10() + 
  labs(y = 'community detection error', 
       colour = NULL) +
  geom_line(aes(x = n, y = med.err,
                colour = 'Algorithm 2')) +
  geom_errorbar(aes(x = n, ymin = first.q, ymax = third.q,
                    colour = 'Algorithm 2')) + 
  geom_line(aes(x = n, y = med.err.ssc,
                colour = 'SSC')) + 
  geom_errorbar(aes(x = n, ymin = first.q.ssc, ymax = third.q.ssc,
                    colour = 'SSC')) + 
  geom_line(aes(x = n, y = med.err.mm,
                colour = 'modularity maximization')) +
  geom_errorbar(aes(x = n, ymin = first.q.mm, ymax = third.q.mm,
                    colour = 'modularity maximization')) +
  scale_colour_brewer(palette = 'Set1') + 
  facet_wrap(~ K, labeller = 'label_both') + 
  scale_x_continuous(breaks = c(256, 512, 1024, 2048, 4096)) +
  scale_y_log10()
```

@@

```{r clust_err_ct_sim, fig.cap = 'IQR of counts of misclustered vertices using Algorithm 2 (red) compared against subspace clustering (blue) for sample sizes from 128 to 4096. Simulations were repeated 40 times for each sample size. The $y$-axis is shown on a logarithmic scale.', cache = FALSE, fig.width = 10}
clustering.df <- readr::read_csv('clustering-k.csv')
clustering.df %>%
  dplyr::group_by(n, K) %>%
  dplyr::summarise(med.err = median(error),
                   first.q = quantile(error, .25),
                   third.q = quantile(error, .75),
                   med.err.ssc = median(error.ssc),
                   first.q.ssc = quantile(error.ssc, .25),
                   third.q.ssc = quantile(error.ssc, .75)) %>%
  dplyr::ungroup() %>% 
  ggplot() +
  # scale_x_log10() + 
  labs(y = 'misclustering count', 
       colour = NULL) +
  geom_line(aes(x = n, y = med.err * n,
                colour = 'Algorithm 2')) +
  geom_errorbar(aes(x = n, ymin = first.q * n, ymax = third.q * n,
                    colour = 'Algorithm 2')) + 
  geom_line(aes(x = n, y = med.err.ssc * n,
                colour = 'SSC')) + 
  geom_errorbar(aes(x = n, ymin = first.q.ssc * n, ymax = third.q.ssc * n,
                    colour = 'SSC')) + 
  scale_colour_brewer(palette = 'Set1') + 
  facet_wrap(~ K, labeller = 'label_both') + 
  scale_y_log10() +
  scale_x_continuous(breaks = c(256, 512, 1024, 2048, 4096))
```

@@

```{r clust_err_k_imba, fig.cap = 'IQR of clustering error using Algorithm 2 (red) compared against subspace clustering (blue) for sample sizes from 128 to 4096 and number of clusters from 2 to 4. Simulations were repeated 50 times for each sample size. The $y$-axis is shown on a logarithmic scale.', fig.width = 10}
clustering.df <- readr::read_csv('clustering-k-imbalanced.csv')
clustering.df %>%
  dplyr::group_by(K, n) %>%
  dplyr::summarise(med.err = median(error),
                   first.q = quantile(error, .25),
                   third.q = quantile(error, .75),
                   med.err.ssc = median(error.ssc),
                   first.q.ssc = quantile(error.ssc, .25),
                   third.q.ssc = quantile(error.ssc, .75)) %>%
  dplyr::ungroup() %>% 
  ggplot() +
  # scale_x_log10() + 
  labs(y = 'community detection error', 
       colour = NULL) +
  geom_line(aes(x = n, y = med.err,
                colour = 'Algorithm 2')) +
  geom_errorbar(aes(x = n, ymin = first.q, ymax = third.q,
                    colour = 'Algorithm 2')) + 
  geom_line(aes(x = n, y = med.err.ssc,
                colour = 'SSC')) + 
  geom_errorbar(aes(x = n, ymin = first.q.ssc, ymax = third.q.ssc,
                    colour = 'SSC')) + 
  scale_colour_brewer(palette = 'Set1') + 
  facet_wrap(~ K, labeller = 'label_both') + 
  scale_x_continuous(breaks = c(256, 512, 1024, 2048, 4096)) +
  scale_y_log10()
```
